A Lyapunov Analysis of Momentum Methods \\ in Optimization

### 2016-10-29 LyapDraft.tex
made several of mike’s suggested changes minus fixing appendix

\begin{document}
\title{A Lyapunov Analysis of Momentum Methods \\ in  Optimization}
%\title{A Lyapunov Perspective on Momentum Methods \\ in  Optimization}
\date{}
\author{Ashia C. Wilson \and Benjamin Recht \and Michael I. Jordan}
\maketitle
\begin{abstract}
Momentum methods play a central role in convex optimization. Several momentum methods are provably optimal and all use a technique called {\em estimate sequences} to analyze their behavior.  The technique of estimate sequences has long been considered difficult to understand. In this paper we show there is an equivalence between the technique of estimate sequences and a family of Lyapunov functions in both continuous and discrete time. This framework allows us to develop a simple and unified analysis of several existing algorithms, introduce a couple of new algorithms, and strengthen the connection between algorithms and continuous time dynamical systems. 
\end{abstract}

@@

In this paper, we propose a bridge between continuous-time limits and discrete-time algorithms.  Our method takes, as its primary object of study a general Lyapunov function for momentum methods in continuous time.  Through a diverse set of examples, we demonstrate how the proofs of momentum methods can be understood as bounding discretization errors of the Lyapunov function when moving to discrete time.  In particular, we show how the discretization of the associated continuous time ODE needs to be performed in such a way that the Lyapunov function remains valid when transitioning to discrete time.
Using this technique, we provide a simple, direct proof of the convergence of Nesterov's method for strongly convex functions in Euclidean spaces.  Our proof makes clear why only certain parameterizations of the continuous ODEs lead to valid discrete-time methods.  Moreover, we explain the need for the extragradient step inside Nesterov's method which does not appear in Polyak's method.  
Finally, we show there is an equivalence between estimate sequences and Lyapunov functions.  In continuous and discrete time, estimate sequences can be derived directly from the Lyapunov function and vice versa.  We show that the associated continuous-time estimate sequence can also be directly discretized to give standard estimate-sequence based proofs of momentum methods.  This clarifies how Nesterov's ``algebraic trick'' is closely related to invariant reduction proofs that are common in algorithm analysis.

@@

\paragraph{Momentum Methods}
The intuition for solving problem~\eqref{eq:main-problem} with momentum methods comes from viewing optimization algorithms as discrete-time approximations to ODEs. In particular, the second-order ODE:
 \begin{align}\label{eq:heavy-ball-ode}
 m \frac{d^2x}{dt^2}& = - \nabla f(x) - m \gamma \frac{dx}{dt}
 \end{align}
corresponds to the equations of motion of an object with mass $m$ in the presence of  friction or viscosity as well as a potential $f$.  Note that the fixed points of such dynamics are those where the gradient of $f$ vanishes.   The trajectories of this dynamical system will tend to have directions of motion that persist over time.  In the limit as the mass goes to zero, and letting $\gamma \propto 1/m$, we recover a gradient-flow equation, which is the continuous-time limit of gradient descent~\cite{Acceleration}.  Thus, the mass term here serves to \emph{accelerate} the progress towards the bottom of the well.

@@

Wibisono, Wilson, and Jordan~\cite{Acceleration} introduced a class of dynamical systems whose discretizations give rise to a general family of accelerated gradient algorithms, including gradient descent and mirror descent. These algorithms are obtained as a discretization of the following second-order ODE:
 \begin{align} \label{Eq:ELBreg} 
\frac{d}{dt} \nabla h(X_t + e^{-\alpha_t} \dot X_t) = -e^{\alpha_t+\beta_t} \nabla f(X_t)\,,
\end{align}
where $h$ is a strongly convex function associated with the geometry of the space and where $\alpha_t$ and $\beta_t$ are smooth functions of time.  In the case where $\alpha_t=-\log m$, and $\beta_t = \log m$,  the continuous-time limit $m \rightarrow 0$ of this ODE recovers the mirror descent dynamics~\cite[(78)]{Acceleration}, with $h$ playing the role of the distance-generating function.  Indeed, the original motivation for mirror descent by Nemirovski and Yudin was from this continuous perspective, and they demonstrated a Lyapunov argument for the convergence of the continuous time dynamics (see Section 3.1 of Nemirovski and Yudin~\cite{NemirovskiiYudin}).

@@

%\begin{small}
\begin{subequations}\label{Eq:LyapAnal}
\begin{align}
 \frac{d}{dt}D_h\left(x, Z_t\right)  &= \frac{d}{dt}\left( h(x) - h(Z_t) - \langle \nabla h(Z_t), x- Z_t\rangle \right) \notag\\
& = - \langle \nabla h(Z_t), \dot Z_t\rangle -\left\langle \frac{d}{dt} \nabla h(Z_t),x - Z_t\right\rangle  + \langle \nabla h(Z_t), \dot Z_t\rangle \notag\\ 
 &= - \left\langle \frac{d}{dt} \nabla h\left(Z_t\right), x - Z_t\right\rangle \, \notag\\%\label{eq:apply-identity}\\
&=    \left\langle \frac{d}{dt}\left(e^{\beta_t}\right)\nabla f(X_t), x  - X_{t} - \frac{1}{\dot \beta_t} \dot X_{s} \right\rangle  \, \label{eq:apply-dynamics}   \\
& =  \frac{d}{dt}\left(e^{\beta_t}\right) \langle \nabla f(X_t), x - X_t \rangle\, - e^{ \beta_t} \langle \nabla f(X_t), \dot X_t\rangle dt\, \notag   \\
& = \frac{d}{dt} \left(e^{\beta_t}\right) \langle \nabla f(X_t), x - X_t \rangle\,  -  \frac{d}{dt} \left(e^{ \beta_t} f(X_t)\right)\,   + \frac{d}{dt} \left(e^{\beta_t}\right)  f(X_t)  \notag\\
&= \frac{d}{dt}\left( e^{\beta_t}\right) [f(X_t)+ \langle \nabla f(X_t), x - X_t \rangle] -\frac{d}{dt} \left(e^{ \beta_t} f(X_t)\right)\,  \notag  \\ %\label{Eq:Line9}\\
& \leq   \frac{d}{dt}\left( e^{\beta_t}\right) f(x) - \frac{d}{dt} \left(e^{ \beta_t} f(X_t)\right)\, \label{eq:apply-convexity}\\
& = - \frac{d}{dt} \left(e^{\beta_t}\left(f(X_t) - f(x)\right)\right) \label{Eq:cond1}.
\end{align}
\end{subequations}
%\end{small}
Here~\eqref{eq:apply-dynamics} uses the momentum dynamics%~\eqref{Eq:EL}.
~\eqref{Eq:ELH} and~\eqref{Eq:ELZ}. 
The inequality~\eqref{eq:apply-convexity} follows from the convexity of $f$. Rearranging terms and taking $x = x^\ast$, we have show that the function~\eqref{Eq:E} has nonpositive derivative for all $t$ and is hence a Lyapunov function for the family of momentum dynamics~\eqref{Eq:EL}.

@@

The subequations corresponding to~\eqref{Eq:AlgoForward1} may be difficult to solve. For reference, the proximal minimization method~\cite{ChenTeboulle93}, 
\begin{equation}
x_{k+1} = \arg\min_{x\in\X} \left\{ f(x) + \frac{1}{A_{k+1} - A_k}D_h\left(x, x_k\right)\right\},
\end{equation}
involves applying the implicit-Euler method to the mirror descent dynamic, and both yield a $O(1/A_k)$ convergence rate. 
%MJ This sentence isn't clear, and I'm not sure it's worth including.
%Nevertheless, the proof structure which we demonstrated from this thought experiment serves as the backbone for analyzing many discretizations of \eqref{Eq:DynaBeta}, several of which constitute popular methods used throughout optimization.

@@

is a Lyapunov function for any dynamic satisfying~\eqref{Eq:Frank1} and~\eqref{Eq:Frank2}.  We now turn our attention to the conditional gradient method, which maintains the following iterates: 

@@

In this section, we introduce a new family of momentum dynamics which can be discretized to obtain the accelerated gradient descent algorithm in the setting where the objective function $f$ is strongly convex. In Appendix~\ref{App:StrongConvexity}, we show how to derive a Lyapunov function using the dynamic. Furthermore, we show how to discretize the dynamic to obtain a discrete-time algorithm which can be analyzed by performing an analogous discretization of the Lyapunov function. 

@@

Choose $A_0 = 1$, $x_{-1}=x_0 = z_0$ and $\tau_{k+1} = \frac{A_{k+1} - A_k}{A_k}$. Define recursively,

@@

Choosing $\tau_k = \sqrt{\kappa}$ gives the $O(e^{-\sqrt{\kappa} k})$ convergence rate. However, there is no restriction on $\tau_k$;  we are free to scale it arbitrarily. Given the subproblems for this algorithm are difficult to solve, we consider other discretizations based on the explicit-Euler method. In particular, we utilize the same trick of introducing a new sequence $\{y_k\}_{k=1}^\infty$ in order to bound the error. We find that we can use a Lyapunov argument to analyze the following two sequences: 

@@

In this section, we connect our Lyapunov framework to the technique of estimate sequences.  We derive continuous time estimate sequences directly from our Lyapunov function and demonstrate how these two techniques are equivalent in discrete time.

@@

The continuous time view shows that the errors in both the Lyapunov function and estimate sequences are due to discretization errors. We provide a sketch for how this works for most of these methods, leaving some of the details for Appendix~\ref{App:EstSeq}.

@@

\clearpage
\appendix
 \section{Mirror Descent Dynamic}
 \label{App:Mirror}
 We begin with a smooth manifold $\X$ with a local metric $\mathrm{g}(x)$ and a function $f:\X \rightarrow \R$ we would like to minimize. The gradient flow associated with $f$ is the flow induced by the differential equation,
\begin{equation}
\dot X_t = v(X_t),
\end{equation} where the vector field $v(X_t)$ is the ``steepest descent'' direction (the direction that makes $f$ decrease the fastest)
\begin{equation}
v(x) = \arg \min _v \left\{ \langle \nabla f(X_t), v\rangle + \frac{1}{2}\|v\|_{\mathrm{g}(X_t)}^2\right\}.
\end{equation}
We can write the gradient flow equation explicitly as,
 %along which $f$ decreases the most, when measured by the metric $\mathrm{g}(X_t)$. If we unpack this statement, notice that  % which characterized the direction of steepest descent. In particular, 
%\begin{equation}
%\frac{d}{dt} f(X_t) = \langle \nabla f(X_t), \dot X_t\rangle = \langle \mathrm{g}(X_t)^{-1} \nabla f(X_t), \dot X_t\rangle_x = \langle \mathrm{g}(X_t)^{-1} \nabla f(X_t), v(X_t)\rangle_x.
%\end{equation} 
%From above, we can conclude that the direction $v(X_t)$, along which $f$ can be said to decrease the most when measured by the metric, is $v(X_t) = -\mathrm{g}(X_t)^{-1}\nabla f(X_t)$.  Thus, for a smooth manifold $\X$ with local metric $\mathrm{g}(x)$, the gradient flow equation can be written as
\begin{equation}\label{Eq:ManGradFlow}
\dot X_t = - \mathrm{g}(X_t)^{-1} \nabla f(X_t).
\end{equation}
%
%the interpretation of a gradient flow as a steepest descent flow becomes more obvious. 
%However, it is often thought of vector field  the metric to map the differential of the function $\nabla f(x)$ to the tangent space. 
%Intuitively, the gradient is a directional derivative and is therefore an element of the cotangent space. The metric is used to map the gradient onto the tangent space to figure out the corresponding vector field.  
%\paragraph{The Mirror Descent Dynamic} 
The dynamic which forms the basis for the mirror descent algorithm arises from a special structure that appears when the metric is chosen to be the Hessian of a strictly convex function $h$:
% $\mathrm{g}(x) = \nabla^2 h(x)$, 
\begin{equation}\label{Eq:NatGrad}
\dot X_t = -[\nabla^2 h(X_t)]^{-1} \nabla f(X_t).
\end{equation}
which we can also write as,
\begin{equation}\label{Eq:MirrorFlow}
\frac{d}{dt} \nabla h(X_t) = - \nabla f(X_t).
\end{equation}
While on the surface, the ability to rewrite \eqref{Eq:NatGrad} as \eqref{Eq:MirrorFlow} appears to be a nice algebraic trick, it is actually a manifestation of the following fact: \begin{center}{\em When $h:\X \rightarrow \R$ is strictly convex and $\mathrm{g}(x)= \nabla^2 h(x)$, there is a special map \\$\phi:\X \rightarrow \mathcal{Y}$ called the {\em mirror map}, given by $\phi = \nabla h$, whose push-forward maps\\ a gradient flow on $\X$ to a gradient flow on $\mathcal{Y}$. \cite{Acceleration}}\end{center} As a quick exercise, we can check that the following holds under the mirror map:

@@

 That is,  the new curve is obtained by traversing the old curve at a new speed of time. If we consider the arbitrary time-dilation function $\tau(t) = e^{\beta(t)}:= e^{\beta_t}$, where $\dot \beta_t >0$, applied to the mirror descent dynamic \eqref{Eq:MirrorFlow}, we obtain the following equation, %that the time-dilated mirror descent dynamic is given by the equation

@@

 However, notice that the primal form~\eqref{Eq:NatGrad} of the mirror descent dynamic allows us to obtain a stronger guarantee,

@@

%Integrating shows~\eqref{Eq:LyapGrad} is a Lyapunov function for the gradient flow dynamic $\dot X_t = -\nabla f(X_t)$ when $f$ is $\mu$-strongly convex.

@@

%%%%%%%%%%%%%%%%
\subsection{Proof of General Quasi-Monotone Method}
\label{App:ProofQuasiGen}
In \ref{App:ProofQuasi} we have set  the dual averaging term $\gamma_k \equiv 1$  for simplicity.
%\noindent The proof of Theorem~\ref{Thm:Quasi} can be found in Appendix.
%Notice, we have set the dual averaging term $\gamma_k \equiv 1$. This is for simplicity of the presentation. 
In its more general form~\eqref{Eq:XSeqQuasi1} can be written as, 
\begin{equation*}
z_k = \arg \min_{z\in \X} \left\{ \sum_{i=1}^k \alpha_i \langle \nabla f(x_i), z\rangle + \gamma_kD_h(z, z_0)\right\} \label{Eq:XSeqQuasi2},
\end{equation*}
which satisfies the optimality condition,
\begin{align}\label{Eq:DualAvgX}
\gamma_{k+1} \nabla h(x_{k+1})  - \gamma_k \nabla h(x_k)  =-(A_{k+1} - A_{k}) \nabla f(x_{k+1}).
\end{align}
For this method, the notion of a ``prox-center'' is essential. We provide an analysis of  this algorithm using the Lyapunov function 
\begin{equation}\label{Eq:Lyap3}
E_k = A_k(f(x_k) - f(x^\ast)) + \gamma_k (D_h(x^\ast, z_k) -D_h(x^\ast, x_0)),
\end{equation}
however first, we analyze the dynamic to understand how one could obtain the Lyapunov function \eqref{Eq:Lyap3} for the dual-averaged algorithm.

>>>>

\begin{document}
\title{A Lyapunov Analysis of Momentum Methods \\ in  Optimization}
%\title{A Lyapunov Perspective on Momentum Methods \\ in  Optimization}
\date{}
\author{Ashia C. Wilson \and Benjamin Recht \and Michael I. Jordan}
\maketitle
\begin{abstract}
Momentum methods play a central role in convex optimization. Several momentum methods are provably optimal and all use a technique called {\em estimate sequences} to analyze their behavior.  The technique of estimate sequences has long been considered difficult to understand. In this paper we show there is an equivalence between the technique of estimate sequences and a family of Lyapunov functions in both continuous and discrete time. This framework allows us to develop a simple and unified analysis of several existing algorithms, introduce a couple of new algorithms, and strengthen the connection between algorithms and continuous-time dynamical systems. 
\end{abstract}

@@

In this paper, we propose a bridge between continuous-time limits and discrete-time algorithms.  Our method takes, as its primary object of study a general Lyapunov function for momentum methods in continuous time.  Through a diverse set of examples, we demonstrate how the proofs of momentum methods can be understood as bounding discretization errors of the Lyapunov function when moving to discrete time.  In particular, we show how the discretization of the associated continuous-time ODE needs to be performed in such a way that the Lyapunov function remains valid when transitioning to discrete time.
Using this technique, we provide a simple, direct proof of the convergence of Nesterov's method for strongly convex functions in Euclidean spaces.  Our proof makes clear why only certain parameterizations of the continuous ODEs lead to valid discrete-time methods.  Moreover, we explain the need for the extragradient step inside Nesterov's method which does not appear in Polyak's method.  
Finally, we show there is an equivalence between estimate sequences and Lyapunov functions.  In continuous and discrete time, estimate sequences can be derived directly from the Lyapunov function and vice versa.  We show that the associated continuous-time estimate sequence can also be directly discretized to give standard estimate-sequence based proofs of momentum methods.  This clarifies how Nesterov's ``algebraic trick'' is closely related to invariant reduction proofs that are common in algorithm analysis.

@@

\paragraph{Momentum methods}
The intuition for solving problem~\eqref{eq:main-problem} with momentum methods comes from viewing optimization algorithms as discrete-time approximations to ODEs. In particular, the second-order ODE:
 \begin{align}\label{eq:heavy-ball-ode}
 m \frac{d^2x}{dt^2}& = - \nabla f(x) - m \gamma \frac{dx}{dt}
 \end{align}
corresponds to the equations of motion of an object with mass $m$ in the presence of  friction or viscosity as well as a potential $f$.  Note that the fixed points of such dynamics are those where the gradient of $f$ vanishes.   The trajectories of this dynamical system will tend to have directions of motion that persist over time.  In the limit as the mass goes to zero, and letting $\gamma \propto 1/m$, we recover a gradient-flow equation, which is the continuous-time limit of gradient descent~\cite{Acceleration}.  Thus, the mass term here serves to \emph{accelerate} the progress towards the bottom of the well.

@@

Wibisono, Wilson, and Jordan~\cite{Acceleration} introduced a class of dynamical systems whose discretizations give rise to a general family of accelerated gradient algorithms, including gradient descent and mirror descent. These algorithms are obtained as a discretization of the following second-order ODE:
 \begin{align} \label{Eq:ELBreg} 
\frac{d}{dt} \nabla h(X_t + e^{-\alpha_t} \dot X_t) = -e^{\alpha_t+\beta_t} \nabla f(X_t)\,,
\end{align}
where $h$ is a strongly convex function associated with the geometry of the space and where $\alpha_t$ and $\beta_t$ are smooth functions of time.  In the case where $\alpha_t=-\log m$, and $\beta_t = \log m$,  the continuous-time limit $m \rightarrow 0$ of this ODE recovers the mirror descent dynamics~\cite[(78)]{Acceleration}, with $h$ playing the role of the distance-generating function.  Indeed, the original motivation for mirror descent by Nemirovski and Yudin was from this continuous perspective, and they demonstrated a Lyapunov argument for the convergence of the continuous-time dynamics (see Section 3.1 of Nemirovski and Yudin~\cite{NemirovskiiYudin}).

@@

%\begin{small}
\begin{subequations}\label{Eq:LyapAnal}
\begin{align}
 \frac{d}{dt}D_h\left(x, Z_t\right)  &= \frac{d}{dt}\left( h(x) - h(Z_t) - \langle \nabla h(Z_t), x- Z_t\rangle\right)  \notag\\
& = - \langle \nabla h(Z_t), \dot Z_t\rangle -\left\langle \frac{d}{dt} \nabla h(Z_t),x - Z_t\right\rangle  + \langle \nabla h(Z_t), \dot Z_t\rangle \notag\\ 
 &= -\left\langle \frac{d}{dt} \nabla h\left(Z_t\right), x - Z_t\right\rangle \, \notag\\%\label{eq:apply-identity}\\
&=   \dot \beta_t e^{\beta_t}\left\langle\nabla f(X_t), x  - X_{t} - \frac{1}{\dot \beta_t} \dot X_{s} \right\rangle  \, \label{eq:apply-dynamics}   \\
& =   \dot \beta_t e^{\beta_t} \langle \nabla f(X_t), x - X_t \rangle\, - e^{ \beta_t} \langle \nabla f(X_t), \dot X_t\rangle dt\, \notag   \\
& = \dot \beta_t e^{\beta_t} \langle \nabla f(X_t), x - X_t \rangle\,  -   \frac{d}{dt}\left\{e^{ \beta_t} f(X_t)\right\}\,   + \dot \beta_t e^{\beta_t}  f(X_t)  \notag\\
&= \dot \beta_t e^{\beta_t} [f(X_t)+ \langle \nabla f(X_t), x - X_t \rangle] -\frac{d}{dt}\left\{e^{ \beta_t} f(X_t)\right\}\,  \notag  \\ %\label{Eq:Line9}\\
& \leq  \dot \beta_t e^{\beta_t}f(x) - \frac{d}{dt}\left\{e^{ \beta_t} f(X_t)\right\}\, \label{eq:apply-convexity}\\
& = - \frac{d}{dt} \left\{e^{\beta_t}\left(f(X_t) - f(x)\right)\right\} \label{Eq:cond1}.
\end{align}
\end{subequations}
%\end{small}
Here~\eqref{eq:apply-dynamics} uses the momentum dynamics%~\eqref{Eq:EL}.
~\eqref{Eq:ELH} and~\eqref{Eq:ELZ}. 
The inequality~\eqref{eq:apply-convexity} follows from the convexity of $f$. Rearranging terms and taking $x = x^\ast$, we have show that the function~\eqref{Eq:E} has nonpositive derivative for all $t$ and is hence a Lyapunov function for the family of momentum dynamics~\eqref{Eq:EL}.

@@

The subequations corresponding to~\eqref{Eq:AlgoForward1} may be difficult to solve. For reference, the proximal minimization method~\cite{ChenTeboulle93}, 
\begin{equation}
x_{k+1} = \arg\min_{x\in\X} \left\{ f(x) + \frac{1}{A_{k+1} - A_k}D_h\left(x, x_k\right)\right\},
\end{equation}
involves applying the implicit Euler method to the mirror descent dynamic, and both yield a $O(1/A_k)$ convergence rate. 
%MJ This sentence isn't clear, and I'm not sure it's worth including.
%Nevertheless, the proof structure which we demonstrated from this thought experiment serves as the backbone for analyzing many discretizations of \eqref{Eq:DynaBeta}, several of which constitute popular methods used throughout optimization.

@@

is a Lyapunov function for any family of dynamics satisfying~\eqref{Eq:Frank1} and~\eqref{Eq:Frank2}.  We now turn our attention to the conditional gradient method, which maintains the following iterates: 

@@

In this section, we introduce a new family of momentum dynamics which can be discretized to obtain the accelerated gradient descent algorithm in the setting where the objective function $f$ is strongly convex. In Appendix~\ref{App:StrongConvexity}, we show how to derive a Lyapunov function using the dynamic. Furthermore, we show how to discretize the dynamics to obtain a discrete-time algorithm which can be analyzed by performing an analogous discretization of the Lyapunov function. 

@@

Choose $A_0 = 1$, $x_{-1}=x_0 = z_0$ and $\tau_{k+1} = \frac{A_{k+1} - A_k}{A_{k+1}}$. Define recursively,

@@

Choosing $\tau_k = \sqrt{\kappa}$ gives the $O(e^{-\sqrt{\kappa} k})$ convergence rate. However, there is no restriction on $\tau_k$;  we are free to scale it arbitrarily. Given the subproblems for this algorithm are difficult to solve, we consider other discretizations based on the explicit Euler method. In particular, we utilize the same trick of introducing a new sequence $\{y_k\}_{k=1}^\infty$ in order to bound the error. We find that we can use a Lyapunov argument to analyze the following two sequences: 

@@

In this section, we connect our Lyapunov framework to the technique of estimate sequences.  We derive continuous-time estimate sequences directly from our Lyapunov function and demonstrate how these two techniques are equivalent in discrete time.

@@

The continuous-time view shows that the errors in both the Lyapunov function and estimate sequences are due to discretization errors. We provide a sketch for how this works for most of these methods, leaving some of the details for Appendix~\ref{App:EstSeq}.

@@

\clearpage
\appendix
 \section{Mirror Descent Dynamic}
 \label{App:Mirror}
 We begin with a smooth manifold $\X$ with a local metric $\mathrm{g}(x)$ and a function $f:\X \rightarrow \R$ we would like to minimize. The gradient flow associated with $f$ is the flow induced by the differential equation,
\begin{equation}
\dot X_t = v(X_t),
\end{equation} where the vector field $v(X_t)$ is the ``steepest descent'' direction (the direction that makes $f$ decrease the fastest)
\begin{equation}
v(x) = \arg \min _v \left\{ \langle \nabla f(X_t), v\rangle + \frac{1}{2}\|v\|_{\mathrm{g}(X_t)}^2\right\}.
\end{equation}
We can write the gradient flow equation explicitly as,
 %along which $f$ decreases the most, when measured by the metric $\mathrm{g}(X_t)$. If we unpack this statement, notice that  % which characterized the direction of steepest descent. In particular, 
%\begin{equation}
%\frac{d}{dt} f(X_t) = \langle \nabla f(X_t), \dot X_t\rangle = \langle \mathrm{g}(X_t)^{-1} \nabla f(X_t), \dot X_t\rangle_x = \langle \mathrm{g}(X_t)^{-1} \nabla f(X_t), v(X_t)\rangle_x.
%\end{equation} 
%From above, we can conclude that the direction $v(X_t)$, along which $f$ can be said to decrease the most when measured by the metric, is $v(X_t) = -\mathrm{g}(X_t)^{-1}\nabla f(X_t)$.  Thus, for a smooth manifold $\X$ with local metric $\mathrm{g}(x)$, the gradient flow equation can be written as
\begin{equation}\label{Eq:ManGradFlow}
\dot X_t = - \mathrm{g}(X_t)^{-1} \nabla f(X_t).
\end{equation}
%
%the interpretation of a gradient flow as a steepest descent flow becomes more obvious. 
%However, it is often thought of vector field  the metric to map the differential of the function $\nabla f(x)$ to the tangent space. 
%Intuitively, the gradient is a directional derivative and is therefore an element of the cotangent space. The metric is used to map the gradient onto the tangent space to figure out the corresponding vector field.  
%\paragraph{The Mirror Descent Dynamic} 
The dynamics which forms the basis for the mirror descent algorithm arises from a special structure that appears when the metric is chosen to be the Hessian of a strictly convex function $h$:
% $\mathrm{g}(x) = \nabla^2 h(x)$, 
\begin{equation}\label{Eq:NatGrad}
\dot X_t = -[\nabla^2 h(X_t)]^{-1} \nabla f(X_t).
\end{equation}
which we can also write as,
\begin{equation}\label{Eq:MirrorFlow}
\frac{d}{dt} \nabla h(X_t) = - \nabla f(X_t).
\end{equation}
While on the surface, the ability to rewrite \eqref{Eq:NatGrad} as \eqref{Eq:MirrorFlow} appears to be a nice algebraic trick, it is actually a manifestation of the following fact: \begin{center}{\em When $h:\X \rightarrow \R$ is strictly convex and $\mathrm{g}(x)= \nabla^2 h(x)$, there is a special map \\$\phi:\X \rightarrow \mathcal{Y}$ called the {\em mirror map}, given by $\phi = \nabla h$, whose push-forward maps\\ a gradient flow on $\X$ to a gradient flow on $\mathcal{Y}$. \cite{Acceleration}}\end{center} As a quick exercise, we can check that the following holds under the mirror map:

@@

 That is,  the new curve is obtained by traversing the old curve at a new speed of time. If we consider the arbitrary time-dilation function $\tau(t) = e^{\beta(t)}:= e^{\beta_t}$, where $\dot \beta_t >0$, applied to the mirror descent dynamics \eqref{Eq:MirrorFlow}, we obtain the following equation, %that the time-dilated mirror descent dynamic is given by the equation

@@

 However, notice that the primal form~\eqref{Eq:NatGrad} of the mirror descent dynamics allows us to obtain a stronger guarantee,

@@

%Integrating shows~\eqref{Eq:LyapGrad} is a Lyapunov function for the gradient flow $\dot X_t = -\nabla f(X_t)$ when $f$ is $\mu$-strongly convex.

@@

%%%%%%%%%%%%%%%%
\subsection{Proof of General Quasi-Monotone Method}
\label{App:ProofQuasiGen}
In \ref{App:ProofQuasi} we have set  the dual averaging term $\gamma_k \equiv 1$  for simplicity.
%\noindent The proof of Theorem~\ref{Thm:Quasi} can be found in Appendix.
%Notice, we have set the dual averaging term $\gamma_k \equiv 1$. This is for simplicity of the presentation. 
In its more general form~\eqref{Eq:XSeqQuasi1} can be written as, 
\begin{equation*}
z_k = \arg \min_{z\in \X} \left\{ \sum_{i=1}^k \alpha_i \langle \nabla f(x_i), z\rangle + \gamma_kD_h(z, z_0)\right\} \label{Eq:XSeqQuasi2},
\end{equation*}
which satisfies the optimality condition,
\begin{align}\label{Eq:DualAvgX}
\gamma_{k+1} \nabla h(x_{k+1})  - \gamma_k \nabla h(x_k)  =-(A_{k+1} - A_{k}) \nabla f(x_{k+1}).
\end{align}
For this method, the notion of a ``prox-center'' is essential. We provide an analysis of  this algorithm using the Lyapunov function 
\begin{equation}\label{Eq:Lyap3}
E_k = A_k(f(x_k) - f(x^\ast)) + \gamma_k (D_h(x^\ast, z_k) -D_h(x^\ast, x_0)),
\end{equation}
however first, we analyze the dynamics to understand how one could obtain the Lyapunov function \eqref{Eq:Lyap3} for the dual-averaged algorithm.

 \paragraph{The Dynamic} We analyze the momentum dynamics~\eqref{Eq:ELBreg} with an additional increasing weighting term $\gamma_0\geq 0$, $\forall t\in \R$ (where $\dot \gamma_t \geq 0$):
 \begin{subequations}\label{Eq:DualAvgDyn}
\begin{align}
\frac{d}{dt}\Big(\gamma_t \nabla h(Z_t)\Big) &= \frac{d}{dt}\Big(e^{\beta_t}\Big)\nabla f(X_t)\\
Z_t &= X_t + \frac{1}{\beta_t} \dot X_t 
\end{align}
\end{subequations}
With this additional weighting term we obtain the following bound, 
\begin{subequations}\label{Eq:Last}
  \begin{align}
 \frac{d}{dt} \left( \gamma_t D_h(x^\ast,Z_t) \right) \, dt
&=\dot \gamma_t [h(x^\ast) - h(Z_t)] - \frac{d}{dt} \left(\gamma_t\right) \langle\nabla h(Z_t), x - Z_t \rangle-\gamma_t\left \langle \frac{d}{dt} \nabla h(Z_t), x - Z_t\right \rangle \notag\\
&= \dot \gamma_t [h(x^\ast) - h(Z_t)] - \left \langle \frac{d}{dt} \left(\gamma_t  \nabla h(Z_t)\right), x^\ast - Z_t \right\rangle \notag\\
&  \overset{\eqref{Eq:DualAvgDyn}}{=}  \gamma_t [h(x^\ast) - h(Z_t)]  +  \frac{d}{dt}\left(e^{\beta_t}\right) \left\langle \nabla f(X_t), x^\ast - X_t - \frac{e^{\beta_t}}{\frac{d}{dt} e^{\beta_t}} \dot X_t\right\rangle \notag\\
& \overset{\eqref{Eq:LyapAnal}}{\leq} \dot \gamma_t [h(x^\ast) - h(Z_t)]   - \frac{d}{dt}\left( e^{\beta_t}\left(f(X_t) - f(x^\ast)\right)\right) 
%& \leq  \int_0^t \dot \gamma_s D_h(x^\ast, Z_0)  -\int_0^t \frac{d}{ds}\left\{ e^{\beta_s}\left(f(X_s) - f(x^\ast)\right)\right\} ds,
\end{align}
\end{subequations}
\paragraph{Prox-Center} Here, we relax the assumption $\X = \R^d$ and now take it to be some bounded and compact set. 
We define $X_0$ to be the ``prox-center,'' of $\X$:
\begin{align*}
X_0 = \arg\min_{x \in \X} \,h(x)
\end{align*}
and without loss of generality, we choose $h(X_0) = 0$, so that $h(x) =D_h(x,X_0) \geq 0, \, \forall x \in \X$.
Using this definition, from~\eqref{Eq:Last} we obtain the following bound:
% \begin{align*}
%\int_0^t \frac{d}{ds} \left\{ \gamma_s D_h(x^\ast,Z_s) \right\} \, ds \leq  \int_0^t \frac{d}{ds}\left\{\gamma_s D_h(x^\ast, X_0)\right\}  -\int_0^t \frac{d}{ds}\left\{ e^{\beta_s}\left(f(X_s) - f(x^\ast)\right)\right\} ds
%\end{align*}
%
\begin{align*}
 \frac{d}{dt}\left( e^{\beta_t}\left(f(X_t) - f(x^\ast)\right)\right) \,\leq  \frac{d}{dt} \left( -\gamma_t D_h(x^\ast,Z_t) + \gamma_t D_h(x^\ast, Z_0) \right) ,
\end{align*}
from which we can conclude that the following function, 
\begin{equation}
\E_t = e^{\beta_t}( f(X_t) - f(x^\ast)) + \gamma_t (D_h(x^\ast, Z_t) - D_h(x^\ast, Z_0)),
\end{equation}
is a Lyapunov function. We also obtain the following convergence rate guarantee, 
\begin{equation}\label{Eq:DualAvgBound}
f(X_t) - f(X_0) \leq \frac{\gamma_t D_h(x^\ast,Z_0)}{e^{\beta_t}}.
\end{equation}
